\addchap{Acknowledgements}
%\addcontentsline{toc}{chapter}{Acknowledgements}

%I'm most grateful, very grateful, I'd like to thank, Special thanks. Offer my
%gratitude.

%Offer my enduring gratitude

%Like to express my deep gratitude

%I thankfully acknowledge the fellowships and grants that support my research
%during the course of my studies.


First and foremost, I want to express my deep gratitude to my supervisor Roger
Tam, without whom I would not have started my PhD at UBC. From the very
beginning of my internship at the MS/MRI research group until the end of my PhD,
Roger has created a very kind, enjoyable, and productive working environment. He
gave me the freedom and trust, which allowed me to pursue my personal research
interests. I will miss our weekly meetings, which not only allowed me to discuss
new ideas and potential struggles, but which were also filled with many joyful
conversations.  I'm most grateful for the unprecedented amount of time he
dedicates to the supervision of his students, which has always motivated me to
do the best research I can.

I am very grateful for the continued support and valuable feedback from my
co-supervisor Rafeef Abugharbieh and all members of my supervisory committee:
Purang Abolmaesumi, Jane Z. Wang,  and Ghassan Hamarneh. I would also like to
thank my departmental and university defense committee, Robert Rohling, Jim
Little, and Martin McKeown, and the external examiner Daniel Rueckert for their
valuable criticism and insightful feedback.

I'd like to thank everyone from the MS/MRI research group I've had the pleasure
to be working with. Special thanks goes to Anthony Traboulsee and David K.B. Li for
providing insightful feedback about potential clinical applications of my
research, Andrew Riddehough for providing a unique perspective about the
operation of clinical trials, Kevin Lam and Rasmus Storjohann, who have taught
me lots about software development, and Ken Bigelow and Vilmos Soti for
maintaining a productive working environment.

I want to thank all my friends I've met at UBC who've made my everyday a
joyful one. I want to thank my labmates Chunfang, Fahime, Saurabh, Youngjin,
Robert, Lisa, Vanessa, and Ni\~no for many stimulating discussions and welcome
distractions, when I needed one. I also want to thank my friends from the other
labs, Al, Antonio, Colin, Nishant and everyone from BMEGSA and ECEGSA for making
me feel a part of the UBC/SFU research community.

Finally, I want to express my gratitude to my family, who have supported me
throughout all these years. I want to thank my parents and my parents-in-law,
for being understanding and supportive of our decision to move to Canada, even
if it meant that we sometimes won't see each other for more than a year.
I also owe the deepest gratitude to my wife Anne, who has just been amazing
since the moment I met her. Her encouragement and warming love has carried me
through all the years and allowed me to find the right balance between focusing
on my research and enjoying this beautiful country.
